---
title: "Meta-analysis SSC"
author: "jean-paul abbuehl"
date: "02/24/2015"
output: html_document
---

Goal:
Combine SSC/MSC datasets and relate each others with our.
Perform GO term analysis

Prerequisite:
raw microarrays were downloaded from GEO, and preprocessed with Aroma Affymetrix and custom arrays definition from BrainArray.
Datasets were saved in ./datasets

1. Load Datasets and perform a pre-analysis check
```{r}
load('dataset/GSE51608.rda') # Ebf2 dataset
load('dataset/GSE21941.rda') # Nestin dataset
load('dataset/GSE43613.rda') # Prx1-CXCL12 dataset
load('dataset/GSE17569.rda') # Alcam dataset
load('dataset/GSE17570.rda') # Pas dataset

# Check datasets with replicates if they cluster altogether
load('dataset/GSE33158.rda') # Scf dataset
plot(hclust(dist(t(GSE33158)))) # ok

load('dataset/GSE57729.rda') # Grem1 dataset
plot(hclust(dist(t(GSE57729)))) # ok

load('dataset/GSE64447.rda') # Many stromal subpopulations dataset
plot(hclust(dist(t(GSE64447)))) #not ok, replicates don't cluster together

load('dataset/GSE79091.rda') # This study's dataset
plot(hclust(dist(t(GSE79091)))) # ok

# Process GSE64447
correlation<-cor(GSE64447)
to_kick_out<- c()
# For preBCSP
best<-apply(correlation[1:3,1:3],1,function(x){n<-length(x); output<-sort(x,partial=n-1)[n-1]; return (output)})
to_kick_out <- append(to_kick_out, names(which.min(best)))
# For BCSP
best<-apply(correlation[4:6,4:6],1,function(x){n<-length(x); output<-sort(x,partial=n-1)[n-1]; return (output)})
to_kick_out <- append(to_kick_out, names(which.min(best)))
# For Thy1
best<-apply(correlation[7:9,7:9],1,function(x){n<-length(x); output<-sort(x,partial=n-1)[n-1]; return (output)})
to_kick_out <- append(to_kick_out, names(which.min(best)))
# For BLSP
best<-apply(correlation[10:12,10:12],1,function(x){n<-length(x); output<-sort(x,partial=n-1)[n-1]; return (output)})
to_kick_out <- append(to_kick_out, names(which.min(best)))
# For 6C3
best<-apply(correlation[13:15,13:15],1,function(x){n<-length(x); output<-sort(x,partial=n-1)[n-1]; return (output)})
to_kick_out <- append(to_kick_out, names(which.min(best)))
# For HEC
best<-apply(correlation[16:18,16:18],1,function(x){n<-length(x); output<-sort(x,partial=n-1)[n-1]; return (output)})
to_kick_out <- append(to_kick_out, names(which.min(best)))

GSE64447 <- GSE64447[,!(colnames(GSE64447) %in% to_kick_out)]
plot(hclust(dist(t(GSE64447)))) # better than before
```

2. Merge datasets and correct for cross studies comparison
```{r}
merge_dataframe<-function(data1,data2){
  #Cbind data2 in data1
  common1<- rownames(data1) %in% rownames(data2)
  common2<- rownames(data2) %in% rownames(data1)
  data1<-data1[common1,]
  data2<-data2[common2,]
  coordinates<-match(rownames(data1),rownames(data2))
  data2<-data2[coordinates,]
  data1<-cbind(data1,data2)
  return (data1)
}

full_data<-GSE17569
full_data<-merge_dataframe(full_data,GSE17570)
full_data<-merge_dataframe(full_data,GSE51608)
full_data<-merge_dataframe(full_data,GSE21941)
full_data<-merge_dataframe(full_data,GSE43613)
full_data<-merge_dataframe(full_data,GSE33158)
full_data<-merge_dataframe(full_data,GSE57729)
full_data<-merge_dataframe(full_data,GSE64447)
full_data<-merge_dataframe(full_data,GSE79091)

plot(hclust(dist(t(full_data))),main='Before correction')

library(sva)
library(limma)
studies_batch <- as.factor(unlist(list(rep(1,ncol(GSE17569)),rep(2,ncol(GSE17570)),rep(3,ncol(GSE51608)), rep(4,ncol(GSE21941)), rep(5,ncol(GSE43613)), rep(6,ncol(GSE33158)), rep(7,ncol(GSE57729)), rep(8,ncol(GSE64447)), rep(9,ncol(GSE79091)))))

full_data_corrected <- ComBat(full_data,batch=studies_batch,mod=NULL)

plot(hclust(dist(t(full_data_corrected))),main='After correction')
```

3. Collapse arrays between best replicates
```{r}
output<-full_data_corrected[,1:6]
colnames(output)<-c('ALCAMneg_SCAneg','ALCAMpos_SCAneg','ALCAMneg_SCApos','PDGFRneg_SCAneg','PDGFRpos_SCAneg','PDGFRpos_SCApos')
output<-cbind(output,apply(full_data_corrected[,7:9],1,median))
colnames(output)[ncol(output)]<-'EBF2pos'
output<-cbind(output,apply(full_data_corrected[,10:12],1,median))
colnames(output)[ncol(output)]<-'EBF2neg'
output<-cbind(output,apply(full_data_corrected[,13:15],1,median))
colnames(output)[ncol(output)]<-'Nestin'
output<-cbind(output,apply(full_data_corrected[,16:17],1,median))
colnames(output)[ncol(output)]<-'CXCL12pos'
output<-cbind(output,apply(full_data_corrected[,19:21],1,median))
colnames(output)[ncol(output)]<-'SCFpos'
output<-cbind(output,apply(full_data_corrected[,22:24],1,median))
colnames(output)[ncol(output)]<-'SCFneg'
output<-cbind(output,apply(full_data_corrected[,25:27],1,median))
colnames(output)[ncol(output)]<-'GREMneg'
output<-cbind(output,apply(full_data_corrected[,28:30],1,median))
colnames(output)[ncol(output)]<-'GREMpos'
output<-cbind(output,apply(full_data_corrected[,31:32],1,median))
colnames(output)[ncol(output)]<-'preBCSP'
output<-cbind(output,apply(full_data_corrected[,33:34],1,median))
colnames(output)[ncol(output)]<-'BCSP'
output<-cbind(output,apply(full_data_corrected[,35:36],1,median))
colnames(output)[ncol(output)]<-'Thy1'
output<-cbind(output,apply(full_data_corrected[,37:38],1,median))
colnames(output)[ncol(output)]<-'BLSP'
output<-cbind(output,apply(full_data_corrected[,39:40],1,median))
colnames(output)[ncol(output)]<-'6C3'
output<-cbind(output,apply(full_data_corrected[,41:42],1,median))
colnames(output)[ncol(output)]<-'HEC'
output<-cbind(output,apply(full_data_corrected[,c(44,46,48)],1,median))
colnames(output)[ncol(output)]<-'CD73posCD105neg'
output<-cbind(output,apply(full_data_corrected[,c(45,47,49)],1,median))
colnames(output)[ncol(output)]<-'CD73posCD105pos'

library(dendextend)
data <-normalizeQuantiles(output)
dend <-as.dendrogram(hclust(dist(t(data[,-c(1,2,3,4,5,6,8,12,13)]))))
dend %>% hang.dendrogram(hang_height=2) %>% set("branches_k_color", k = 4) %>% set("branches_lwd", 4) %>% plot(horiz = TRUE)
save(output,file='dataset/full_dataset.rda')
```

4. Go term analysis
```{r}
load('dataset/full_dataset.rda')

library(biomaRt)
# select mart and data set
# If dataset/GO_ensembl_musculus.rda is missing, run this code to download it (it may take a while)
# bm <- useMart("ensembl")
# bm <- useDataset("mmusculus_gene_ensembl", mart=bm)
# 
# # Get ensembl gene ids and GO terms
# EG2GO <- getBM(mart=bm, attributes=c('ensembl_gene_id','namespace_1003','go_id','name_1006'))
# EG2GO <- EG2GO[EG2GO$go_id != '',]
load('dataset/GO_ensembl_musculus.rda')

library(limma)
# Perform Go analysis between two branches
data <- normalizeQuantiles(output[,c(7,9,10,11,14,15,16,17,18,19,20,21,22)])
design<-cbind(rep(1,13),c(1,0,0,1,1,1,1,0,0,0,0,1,0))

fit<-lmFit(data,design)
fit<-eBayes(fit)
results<-topTable(fit,coef=2,number=nrow(data),p.value=1)

up<-results[results$logFC>0,]
down<-results[results$logFC<0,]

library(topGO)

geneID2GO <- by(EG2GO$go_id, EG2GO$ensembl_gene_id, function(x) as.character(x))
GO2geneID <- by(EG2GO$ensembl_gene_id, EG2GO$go_id, function(x) as.character(x))

### Build GO data analysis
up_genelist<-up$adj.P.Val
names(up_genelist)<-gsub('_at','',rownames(up))

down_genelist<-down$adj.P.Val
names(down_genelist)<-gsub('_at','',rownames(down))

topDiffGenes_05 <- function(allScore) {return(allScore < 0.05)}

GOdata_up05 <- new("topGOdata",
              description = "SSC vs committed niche cells, pval<0.05",
              ontology = "BP",
              allGenes = up_genelist,
              geneSel = topDiffGenes_05,
              annot = annFUN.gene2GO,
              gene2GO=geneID2GO,
              nodeSize = 5)

GOdata_down05 <- new("topGOdata",
              description = "committed niche cells vs SSC, pval<0.05",
              ontology = "BP",
              allGenes = down_genelist,
              geneSel = topDiffGenes_05,
              annot = annFUN.gene2GO,
              gene2GO=geneID2GO,
              nodeSize = 5)

resultKS_up05 <- runTest(GOdata_up05, algorithm = "weight01", statistic = "ks")
resultKS_down05 <- runTest(GOdata_down05, algorithm = "weight01", statistic = "ks")

allRes_up05 <- GenTable(GOdata_up05, weightKS = resultKS_up05,orderBy = "weightKS", ranksOf = "weightKS", topNodes = 10)
allRes_down05 <- GenTable(GOdata_down05, weightKS = resultKS_down05,orderBy = "weightKS", ranksOf = "weightKS", topNodes = 10)

write.csv(allRes_up05,file='GO term in SSC pval05.csv')
write.csv(allRes_down05,file='GO term in non-SSC pval05.csv')

library(gridExtra)
library(lattice)

# Make graph for upregulated GO term in SSC
pdf('top10 GO term in SSC.pdf',height=20,width=8)
fig<-list()
for(i in 1:nrow(allRes_up05)){
  goID<-allRes_up05[i,'GO.ID']
  fig<-append(fig,list(showGroupDensity(GOdata_up05, goID, ranks = TRUE)))
}

grid.arrange(fig[[1]],fig[[2]],fig[[3]],fig[[4]],fig[[5]],
             fig[[6]],fig[[7]],fig[[8]],fig[[9]],fig[[10]],
             ncol=2)
dev.off()

# Make graph for selected upregulated GO term in non-SSC (there are too many)
goID<-c("GO:0048661","GO:0030097","GO:0032496","GO:0030183","GO:0002544","GO:0071354","GO:0045087")
pdf('top10 GO term in non-SSC.pdf',height=20,width=8)
fig<-list()
for(i in 1:length(goID)){
  id<-goID[i]
  fig<-append(fig,list(showGroupDensity(GOdata_down05, id, ranks = TRUE)))
}

grid.arrange(fig[[1]],fig[[2]],fig[[3]],fig[[4]],fig[[5]],
             fig[[6]],fig[[7]],
             ncol=2)
dev.off()
```